---
title: "R Session I - Working with Files"
output: 
  html_document: 
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Working Directory

When you read or write files, there is a default location which is called the **'working directory'**.  

**getwd()** shows where your working directory will be.



```{r}
getwd()

```


If you want to change that directory use **setwd()**.  

```{r}
#setwd ("c:/R")
```


## Reading and Writing Files

If you want to save your data sets you have a few options.  

### R Objects

Any thing in memory can be stored as a R object to file.   It will preserve the exact formating and content of the file.  Of course, all of this will happen in your working directory.

**save()** will write the R object.
**load()** will load the R object into memory.

```{r}
data (iris)
save (iris,file = "iris.r")
load ("iris.r")
```

When you want to save outputs of models or other odd objects, you will need to use this method.   


### CSV (or other) Files

For things like dataframes, you often want to write to standard files like CSV.   The biggest advantage is that this file format can be read/used by other programs.

```{r}
write.csv (iris, file = "iris.csv", row.names = FALSE)
iris <- read.csv ("iris.csv")
```


## Working with Databases

Ideally you will be reading your data from a database.  For almost any type of database, you will find a connector to it.   We want go through the details because it will depend on the database you are using.  You will likely need a database driver installed on your machine with the right package.

For best practice, you should do as much of your work on the database and limit R to only what you can't do on the database.

**Packages:**  
1. RODBC  
2. RJDBC  
3. RMySQL   
4. RPostgreSQL   

