---
title: "R Session I - Logic, Comparisons and NA"
date: "June 4, 2017"
output: 
  html_document: 
  theme: cosmo
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Logic and Comparisons

A comparison between two values results in either a **TRUE** or a **FALSE** value being returned.

```{r}
#Is 1 equal to 1?
1 == 1

```



1. **==** equal to 
2. **!=** not equal to
3. **>** greater than
4. **<** less than
5. **|** OR
6. **&** AND
7. **!** Not
8. **%in%** In - compares set

```{r}
#Does 1 equal 1 and 2 = 1
(1==1)&(2==1)

```


So, how does this work with vectors? 

1. Remember that when you compare two vectors, you compare each element at a time.
2. If you compare a single value to a vector, it takes that single value and compares it to each element in the vector.

```{r}
a <- c(1,2,3,4)
b <- c(1,2,3,4)
a == b
```

```{r}

a == 1
```

Now let's compare using **%in%**.  It compares the items on the left and returns a TRUE/FALSE if they exist in the set on the right.

```{r}

c(1,2) %in% a

```



##NA Values

When a data set has a value that is missing (i.e. no value exists), it is **NA**.  Depending on the task in R, you will need to replace these with real values.   Some functions will treat NA by ignoring them.

1. **is.na()** will check a vector for any NA values returning a TRUE/FALSE vector
2. **na.omit()** will remove rows that contain NA
3. **complete.cases()** will check a dataframe row-by-row for NA values returning a TRUE/FALSE vector.


```{r}
a <- c(NA, 1, 2, 3)
is.na (a)

```

When we do calculations, we need to determine how to treat the values.  When we average a number, how should we treat NA?  Do we ignore them or do we return a signal that it failed?


```{r}
mean (a)

```

Some functions will have a parameter (an optional switch), that allows you to ignore them.

```{r}
mean (a, na.rm=TRUE)

```

##Square Bracket Operator and Logical Vectors

Ok, so we know we can use the **[]** to subset a vector or dataframe.  We mostly used numeric positions to indicate what we wanted or didn't want.  However, the far more powerful thing is to use TRUE/FALSE vectors.

```{r}
#Let's remove all values from 'a' that are NA
#Since is.na() returns TRUE if NA, we want to reverse that so we use the '!' to reverse the TRUE and FALSE

a[!is.na(a)]

```

The above method can be used to filter data or check any type of condition you want.

```{r}
a [a > 2]

```

Here we check whether the values of 'a' are greater than 2, if they are we will retrieve them.