---
title: "R Session I - Exercise"
output:   
  html_document: 
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Part 1

1. First map your local directory to something you prefer using setwd().
2. Load the CSV into R and save it to a dataframe named “springleaf”
3. How many rows and columns are there? What’s the total cell count?
4. Run the summary() function on the first 10 columns of the dataset
5. Create a new variable (separate from the springleaf dataframe) that stores the length of the ID variable in the springleaf dataframe


```{r}

#setwd(“c:/”) - I won't use it since I'm on a linux server
springleaf <- read.csv (file = "springleaf_small.csv")
nrow(springleaf)
ncol(springleaf)
nrow(springleaf)*ncol(springleaf)
summary (springleaf[,1:10])
ids <- length (springleaf$ID)

```

## Part 2

1. Look up the complete.cases() function in help and read what it does.
2. Now use the complete case function to generate a logical vector. Now try to pass that vector into the function table () which we will cover soon.   What would you conclude about the data set?
3. Now check the variable VAR_0237… what type is it? Now, determine the proportion of states that are either CA, NY or NJ?
4. Let’s add a TRUE / FALSE variable for whether the state is in the CA/NY/NJ list to our exising springleaf dataframe


```{r}

complete.cases (springleaf)
table (complete.cases (springleaf))
str (springleaf$VAR_0237)
table ((springleaf$VAR_0237) %in% c("CA", "NY", "NJ"))
springleaf$VAR_0237 %in% c("CA", "NY", "NJ")
springleaf$IsLiberal <- springleaf$VAR_0237 %in% c("CA", "NY", "NJ")


```

## Part 3

1. Try running table (is.na()) on VAR_0207, VAR_0213 and VAR_0840
2. Now let’s remove these 3 columns.   One simple way is to use:  springleaf$VAR_0207 <- NULL
3. Try running complete.cases() again without those 3 variables
4. Next let’s try to run the mean VAR_0006.
5. Fix the issue by using the parameters in the mean function.




```{r}


table(is.na (springleaf$VAR_0207))
table(is.na (springleaf$VAR_0213))
table(is.na (springleaf$VAR_0840))

springleaf$VAR_0207 <- NULL
springleaf$VAR_0213 <- NULL
springleaf$VAR_0840 <- NULL

mean (springleaf$VAR_0006)
mean (springleaf$VAR_0006, na.rm = TRUE)


```


## Part 4

1. Run the mean, median and quartile functions for VAR_0583.
2. Do you see an issue with the data?
3. Can you fix the issue and create a new cleaned variable?

```{r}

mean (springleaf$VAR_0583)
median (springleaf$VAR_0583)
quantile (springleaf$VAR_0583)

a <- springleaf$VAR_0583
table (cut (a, c(-Inf, mean (a) - 1.5*sd(a), mean(a)+1.5*sd(a), Inf)))
springleaf$VAR_0583 [springleaf$VAR_0583 > 8.91e+08] <- NA
quantile (springleaf$VAR_0583, na.rm = TRUE)

```

## Part 5

1. Let’s create two data sets, one for training (75%) and one for validation (25%).
2. Use table to validate that the target variable proportion is consistent between the two data sets. 
3. Validate that all IDs have been allocated to either test or validation


```{r}

validationIDs <- sample (springleaf$ID, nrow(springleaf)*.75)
validation <- springleaf[springleaf$ID %in% validationIDs,]
test <- springleaf[!(springleaf$ID %in% validationIDs),]

prop.table (table (validation$target))
prop.table (table (test$target))

table (springleaf$ID %in% c(test$ID, validation$ID))


```




