---
title: "R Session I - Vectors and Dataframes"
output: 
  html_document: 
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Defining and Creating Vectors

A variable is a single data point - it could be a character or a numerical value.  A vector is a set of variables of the same type. In other programming languages you may have referred to them as arrays.  They are created by the function **c()** - combine.

```{r}

numbers <- c(1,2,3)
characters <- c("a", "b", "c")

```

```{r}
characters
```

```{r}
numbers
```


## Using Vectors

With a vector, you can apply functions against them.  This will work on each item in the vector and create a new vector.  For example:

```{r}
numbers*2
```

You can also have two vectors and do operations between them.   The below code will take each item and multiply it by the item in the same position.   
* 1*1
* 2*2
* 3*3

```{r}
numbers*numbers
```

You can also create vectors easily using the **:** operator.  It will generate a vector from one value to another value.  Note that I don't need to use the **c()** operator as it knows that it's a vector.

```{r}
1:10

```


## Dataframes

A dataframe is the most important type of data structure you'll work with in R.   A dataframe is similar to a table in a database, a SAS dataset, or an excel sheet.  A data frame has columns and rows.   A column in the dataframe must have the same datatype but different columns can have different data types.   This allows us to mix characters, numerics, logical vectors together into a single table.


``` {r}

df <- data.frame (numbers, characters)
df
str (df)

```

Note the use of the function **str()** which shows the structure of the data set.   It's very useful.

A few other helpful functions:

1. **nrow()** - shows the number of rows in a dataframe
2. **ncol()** - shows the number of columns in a dataframe
3. **summary()** - shows the variables in the dataframe and their distribution/contents

In order to manipulate the dataframe you will want to be able to modify parts of it.  Either a column, a row or a specific cell.  

### The $ operator

Accessing a column can be done easily by using the **$** operator (e.g. dataframe$column_name).  

```{r}
df$numbers
```

You can also create a new column using the **$** operator.

```{r}
df$sq_numbers <- df$numbers^2
str (df)
```

### The [] operator

The 'square bracket' operator allows you to access any part of a vector or a dataframe.  It works like this: dataframe [row,column].   

First, lets start by using the [] operator on a vector.

```{r}
#Gives the 2nd element of the vector
numbers[2]

```

Now, let's use this on a dataframe.

```{r}
#Gives the value in the first column of the first row.
df[1,1] 

#Gives the first row
df [1,]

#Gives the first column
df[,1]


```


What if we want several columns?  We can pass a range of column values instead of just a single value.  

```{r}

#This gives us the first two rows of the data
df [1:2,]

```

We can also pass the names of the columns.  **colnames()** provides the name of the columns to us.  

```{r}

#This will pass only columns that match the names we provide 
df [,c("numbers", "characters")]

```
