---
title: "Using R Markdown"
output: 
  html_document: 
    theme: cosmo
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What and Why Markdown?

Markdown allows you to write more than code.   It takes a basic script and makes it a pdf report, a html webpage or simply allows for more interactive coding.  A simple use case is an analysis that you routinely create - it can be created easily as a markdown document and then updated on live data.
 

For a full guide you can see a simple cheatsheet guide [here](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf).   There's a full write up [here](http://rmarkdown.rstudio.com/articles_intro.html).

## Basics

The basics of markdown are simple.   The following below establishes a block of R code.

```{r}

```


## Basic Formating

1. headers - use # to create a header
2. bolding - you can bold text by wrapping it in double asterisks like **this**
3. links can be embeddded by using formating like this [link](rstudio.com)
4. bullets and numbered lists - you can use single asterisks for bullets or numbers with a period

## Inline Coding

You can also insert code with dynamic data which is great for reporting   For instance, Pi is equal to `r 3+.14`.

## Executing Code

You can execute code in a variety of ways:

1. Line by line.   Simply put your cursor on the line you want to execute and you can hit **CTRL + ENTER**
2. A block at a time.   You can execute a block of code by hitting the 'play' button in the top right of the window.
3. The full script.  You can execute the full script by hitting the 'knit' button at the top.   This will compile all code and create the document.