---
title: "Plotting in R Using ggplot2"
subtitle: "R Tutorial II"
output: 
  pdf_document: 
    theme: cosmo
---
***
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


``` {r include = FALSE}
library (ggplot2)
library (dplyr)
data ("mpg")

```

##Why ggplot2?

ggplot2 is the definitive graphing library in R. In almost all cases it's better than the default **plot()** function in R. 



1. Graphs are built in layers that allow tons of flexibility
2. Clean, easy to read graphs 'out of the box'
3. Every possible graph form none to mankind
4. ggplot2 graphs are objects that can be stored
5. Themes are available to allow full customization

##Cheatsheet

There's a fantastic cheatsheet available for ggplot2 which you can find here.  [link](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)

##Layers

In ggplot2, graphs are built in layers.  

**1. Map Data to Graphing Attributes** 

You specify the dataframe you want to graph and use **aes()** to map the data onto aspects of your graph:

* x, y on the axis
* colour
* alpha
* linetype
* size
* weight

```{r}
p <- ggplot (mpg, aes (x= displ, y = hwy, colour = factor (year)))
p
```

**2. Add Geom(s) to Your Graph**

There are dozens but here are a few key ones:  
* geom_histogram  
* geom_bar  
* geom_line  
* geom_point - *a basic scatter point plot*  
* geom_jitter - *a scatter plot with randomization to better show density*  
* geom_boxplot - *a box and whisker plot*  
* geom_density - *a great plot to show the distribution of data*  

```{r}
p+geom_point()
```


**3. Add Titles, Lables and Legends**

xx 

```{r}
p+geom_point()+ggtitle("Highway Mileage and Displacement")+
  ylab("Highway (MPG)")+
  xlab ("Engine Displacement (L)")
```


**4. Apply Themes to Your Graphs**

You can easily override the basic attributes of a graph using themes.  In fact, there's a whole package that provides even more themes.  [link](https://github.com/jrnold/ggthemes)

* theme_fivethirtyeight - *based on the great fivethirtyeight.com site*
* theme_economist - *based on the economist magazine*
* theme_bw - *simple black and white*

```{r}
library (ggthemes)
p+geom_point()+ggtitle("Highway Mileage and Displacement")+
  ylab("Highway (MPG)")+
  xlab ("Engine Displacement (L)")+
  theme_economist()
```

**5. Apply More Geoms**

You can simply add more geoms on top of the base layers - as many as you want.

```{r}
library (ggthemes)
p+geom_point()+ggtitle("Highway Mileage and Displacement")+
  ylab("Highway (MPG)")+
  xlab ("Engine Displacement (L)")+
  theme_economist()+
  geom_smooth()
```


**6. Add even more geoms with other data sets**

You can simply add more geoms on top of the base layers - as many as you want.

```{r}
library (dplyr)
mpg_grouped <- mpg %>% 
  group_by(manufacturer) %>% 
  summarise (mean_hwy = mean(hwy))

Top_Bot <- rbind(mpg_grouped %>% top_n(-1), mpg_grouped %>% top_n(1))

p+geom_point(alpha = .25)+
  ggtitle("Highway Mileage and Displacement")+
  ylab("Highway (MPG)")+
  xlab ("Engine Displacement (L)")+
  theme_economist()+
  geom_smooth()+
  geom_hline(data =Top_Bot, aes (yintercept =mean_hwy))+
  geom_label (aes (label = Top_Bot$manufacturer[1], y = Top_Bot$mean_hwy[1], x = 2), color = "black")+
  geom_label (aes (label = Top_Bot$manufacturer[2], y = Top_Bot$mean_hwy[2], x = 2), color = "black")

```

##Links and Tips

[link](http://zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/)