---
title: "R Session III - Time Series"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Time Series 

``` {r}
require  (forecast)
data (gas)


```


Time series data can be considered to have 3 parts:

1. **Seasonal** - this could be daily, weekly, quarterly or monthly 
2. **Trend** - a longer term movement in the data
3. **Random** - remaining noise that doesn't fit into the above two

Time series analysis takes no other information into account.  So, things like weather, economic factors, etc. that would impact your data won't be factored.

The **decompose** function will take a time series and break it into these components.

``` {r}

plot (decompose(gas))

```


The **forecast** function is the predict counter-part.   If it doesn't have a model, it wil run a model for you using exponential triple smoothing.

**forecast (data, h, level, model)**
**data** - the series we were forecasted
**h** - how many periods in the future to forecast
**level** - the range for the confidence intervals (defaults to 80% and 95%)
**model** - tif you want to specify the model

``` {r}

f <- forecast (gas, h=3, level=.9)
f

arima <- auto.arima(gas)

f2 <- forecast (gas, h= 3, level = .9, model = arima)
f2

```

We can also plot the forecasts pretty easily

``` {r}

plot(f)

```

```{r}
summary (f)
summary (f2)

```